---
title: "Murchieona"
author: "Cenaida Largo"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    number-sections: true
editor: visual
---

# Set up

```{r}
#| label: setup-packages
#| warning: false
#| message: false
library(readr)    # read CSV
library(dplyr)    # data manipulation
library(leaflet)  # map visualization
library(ggplot2)  # data visualization
library(stringr)  # text manipulation
```

# Data import

```{r}
#| label: datasets
#| message: false
taxonomy <- read_csv("data/VERS2022_taxonomy.csv")
assemblages <- read_csv("data/VERS2022_assemblages.csv")
stations <- read_csv("data/VERS2022_stations.csv")
```

# Taxa

Taxa code extraction from the taxonomy file.

```{r}
#| label: clean-taxonomy
taxonomy <- taxonomy %>%
  setNames(as.character(.[2, ])) %>%
  slice(-2)
```

Check     abbreviated_revised_name,
    revised_name,
    full_revised_name,
    justification_name_revision
```{r}
#| label: extract-target-taxa
MUR_MUL <- taxonomy |>
  filter(revised_name == "Murchieona muldali")

MUR_MIN <- taxonomy |>
  filter(revised_name == "Murchieona minuscula")

MUR_MUL_cod <- unique(MUR_MUL$abbreviated_revised_name)
MUR_MIN_cod <- unique(MUR_MIN$abbreviated_revised_name)
```

# Distribution

Useful data from the stations file
```{r}
#| label: stations-selected

stations_selected <- stations |>
  select(
    id_station,
    source,
    latitude_wgs84,
    longitude_wgs84,
    day,
    month,
    corine_land_cover,
    corine_land_cover_particularity,
    geology
  )
```


Occurrences extraction from the assemblages file 

```{r}
#| label: mur-mul-occurrences
MUR_MUL_occ <- assemblages |>
  filter(abbreviated_revised_name == MUR_MUL_cod) |>
  select(
    id_station,
    source,
    abbreviated_revised_name,
    development_stage
  ) |>
  left_join(stations_selected,
            by = c("id_station", "source"))
```

```{r}
#| label: mur-min-occurrences
MUR_MIN_occ <- assemblages |>
  filter(abbreviated_revised_name == MUR_MIN_cod) |>
  select(
    id_station,
    source,
    abbreviated_revised_name,
    development_stage
  ) |>
  left_join(stations_selected,
            by = c("id_station", "source"))
```


```{r}
#| label: combined-distribution
leaflet() %>%
  addTiles() %>%
  addCircleMarkers(
    data = MUR_MIN_occ,
    lng = ~longitude_wgs84,
    lat = ~latitude_wgs84,
    popup = ~id_station,
    radius = 5,
    color = "blue"
  ) %>%
  addCircleMarkers(
    data = MUR_MUL_occ,
    lng = ~longitude_wgs84,
    lat = ~latitude_wgs84,
    popup = ~id_station,
    radius = 5,
    color = "red"
  )
```
Red points correspond to *Murchieona muldali* and blue points to *Murchieona minuscula*.

# Phenology

Records of *M. muldali* by month.

```{r}
#| label: add-month-name-factor
MUR_MUL_occ <- MUR_MUL_occ |>
  mutate(
    month_name = factor(
      month.abb[month],
      levels = month.abb
    )
  )
```

```{r}
#| label: barplot-records
ggplot(MUR_MUL_occ, aes(x = month_name)) +
  geom_bar() +
  labs(
    title = paste("Monthly observations of", MUR_MUL_cod),
    x = "Month",
    y = "Number of observations"
  ) +
  theme_minimal()
```

Development stages of *M. muldali* by month.

```{r}
#| label: barplot-development-stages
ggplot(MUR_MUL_occ, aes(x = month_name, fill = development_stage)) +
  geom_bar(position = "dodge") +
  labs(
    title = paste("Monthly observations of", MUR_MUL_cod),
    x = "Month",
    y = "Number of observations",
    fill = "Development stage"
  ) +
  theme_minimal()
```

Joining adults and subadults.

```{r}
#| label: join-adult-subadult
MUR_MUL_occ <- MUR_MUL_occ %>%
  mutate(
    stage_group = case_when(
      development_stage %in% c("adult", "subadult") ~ "adult/subadult",
      TRUE ~ as.character(development_stage)
    )
  )
```

```{r}
#| label: barplot-joined
ggplot(MUR_MUL_occ, aes(x = month_name, fill = stage_group)) +
  geom_bar(position = "dodge") +
  labs(
    title = paste("Monthly observations of", MUR_MUL_cod),
    x = "Month",
    y = "Number of observations",
    fill = "Development stage"
  ) +
  theme_minimal()
```

# Habitat

using Corine Land Cover classes

```{r}
#| label: CLC-MUR-MUL
MUR_MUL_occ %>% count(corine_land_cover, sort = TRUE)
```

```{r}
#| label: CLC-MUR-MIN
MUR_MIN_occ %>% count(corine_land_cover, sort = TRUE)
```

1.  pastures (231)
2.  broad-leaved forest (311)
3.  natural grassland (321)
4.  inland marshes (411)
5.  fruit trees and berry plantations (222)
6.  mixed forest (313)
7.  transitional woodland shrub (324)
8.  green urban areas (141)
9.  moors and heathland (322)
10. sclerophyllous vegetation (323)

## Are there any records of *M. muldali* in olive groves?

### Agricultural areas

corine_land_cover = 2XX

```{r}
#| label: agri
MUR_MUL_occ_agri <- MUR_MUL_occ %>% 
  filter(str_detect(corine_land_cover, "^2\\d{2}$"))
```

There are `r nrow(MUR_MUL_occ_agri)` records of *M. muldali* in agricultural areas.

### Permanent crops

corine_land_cover = 22X

```{r}
#| label: per
MUR_MUL_occ_per <- MUR_MUL_occ %>% 
  filter(str_detect(as.character(corine_land_cover), "^22\\d$"))
```

There are `r nrow(MUR_MUL_occ_per)` records of *M. muldali* in agricultural areas.

### Olive grooves

corine_land_cover = 223

```{r}
#| label: oli
MUR_MUL_occ_oli <- MUR_MUL_occ %>% 
  filter(corine_land_cover == 223)
```

There are `r nrow(MUR_MUL_occ_oli)` records of *M. muldali* in agricultural areas.

# Geology

Geology classes

```{r}
#| label: geology-count
MUR_MUL_occ %>% count(geology, sort = TRUE)
```

